Optimal updating factor in Wang-Landau and metadynamics simulations

### 2016-01-20 doc/invt.tex
add langtest.c
invt.tex: add discussion on the correlation integrals

To proceed, we first observe that
the average $\bar v = \sum_{i = 1}^n p_i \, v_i$
increases steadily over time:
%
\begin{equation}
\frac{ d \bar v } { d t }
=
\sum_{i = 1}^n p_i \dot v_i
=
\alpha(t) \sum_{i = 1}^n h_i(t) = \alpha(t).
\label{eq:dvbardt}
\end{equation}
%
However, we expect the difference between $v_i$ and $\bar v$,
%
\begin{equation}
  x_i \equiv v_i - \bar v = v_i - \sum_{j = 1}^n p_j \, v_j,
  \label{eq:x_def}
\end{equation}
%
to be small in the asymptotic regime,
and expand Eq. \eqref{eq:h_ave} as
\begin{align}
\langle h_i(t) \rangle
&\approx
\frac{ p_i \, e^{- x_i} }
{ \sum_{ j = 1}^n p_j \, e^{- x_j} }
\approx
\frac{ p_i ( 1 - x_i ) }
{ \sum_{ j = 1}^n p_j (1 - x_j) }
\notag \\
&\approx
p_i \, \left(
  1 - x_i + \sum_{j=1}^n p_j \, x_j
\right)
=
p_i \, (1 - x_i),
\label{eq:hdet_linearize}
\end{align}
where we have used $\sum_{j=1}^n p_j = 1$,
and
%
\begin{equation}
  \sum_{i = 1}^n p_i \, x_i = 0,
  \label{eq:px_sum}
\end{equation}
which follows directly from Eq. \eqref{eq:x_def}.
%
From Eqs.
\eqref{eq:vt_diffeq},
\eqref{eq:h_split},
\eqref{eq:dvbardt}-\eqref{eq:hdet_linearize},
we get a set of decoupled equations
%
\begin{equation}
  \dot x_i(t)
  =
  -\alpha(t) \, \left[ x_i(t) - \frac{ \zeta_i(t) } { p_i } \right].
  \label{eq:dxdt_WL}
\end{equation}
\note{The derivation:
$$
\begin{aligned}
  \dot x_i
  &\stackrel{\mathrm{Eq.\; \eqref{eq:x_def}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=}
  \dot v_i - \dot{ \bar v }
  \stackrel{\mathrm{Eqs.\; \eqref{eq:vt_diffeq}\; and\; \eqref{eq:dvbardt}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left( \frac{ h_i } { p_i } - 1 \right)
  \\
  &\stackrel{\mathrm{Eq.\; \eqref{eq:h_split}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left( \frac{ \langle h_i \rangle } { p_i } + \frac{ \zeta_i } { p_i } - 1 \right)
  \stackrel{\mathrm{Eq.\; \eqref{eq:hdet_linearize}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left[ -x_i(t) + \frac{ \zeta_i } { p_i } \right].
\end{aligned}
$$
}

@@

Further, by Eq. \eqref{eq:w_detailedbalance},
we can define a symmetric matrix $\hat{\mathbf w}$
as $\hat w_{ij} = \sqrt{p_i/p_j} \, w_{ij}$
that can be diagonalized\cite{vankampen}
with a set of orthonormal eigenvectors:
%
$\sum_{i = 1}^n \varphi_{ki} \, \hat w_{ij} = \lambda_k \, \varphi_{kj}$.
%
Thus,
in terms of diagonalizing $\mathbf w$, we have
\begin{equation}
  \sum_{i = 1}^n \phi_{ki} \, w_{ij} = \lambda_k \, \phi_{kj},
  \label{eq:eig_w}
\end{equation}
where
$\phi_{ij} = \sqrt{p_j} \, \varphi_{ij}$ satisfies
the orthonormal conditions\cite{vankampen}:
%
\begin{align}
\sum_{k = 1}^n \phi_{ki} \, \phi_{kj}
&= \delta_{ij} \, p_i,
\label{eq:eig_orthonormal_cols}
\\
\sum_{k = 1}^n \frac{ \phi_{ik} \, \phi_{jk} }{ p_k }
&= \delta_{ij}.
\label{eq:eig_orthonormal_rows}
\end{align}

@@

In the asymptotic regime,
we can approximate $\kappa_i(t)$ as
%
\begin{equation}
  \kappa_i(t) \approx \Gamma_i \, \delta(t)
  \label{eq:kappa_delta}
\end{equation}
%
where, $\Gamma_i = \int_{-\infty}^\infty \kappa_i(t) \, dt$.
%
Then, Eq. \eqref{eq:error_mbin} is simplified as
%
\begin{align}
  \mathcal E
  &=
  \int_0^t
  \sum_{i = 2}^n
  \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr) \, dt'.
  \label{eq:error_mbin1}
\end{align}
%
The next order correction to Eq. \eqref{eq:kappa_delta}
would lead to a correction to $\mathcal E$
that is about $\alpha_{\max}^2$ times as large.

\note{This is equivalent to saying that the Fourier transform
  $$
  \tilde \kappa_i(\omega) = \Gamma_i
  $$
  is a roughly constant.
  %
  For the next order correction we would have
  $$
  \tilde \kappa_i(\omega) = \Gamma_i + \Gamma^{(2)}_i \omega^2 + \dots,
  $$
  which leads to a correction to $\mathcal E$
  $$
  \Gamma^{(2)}_i
  \int_0^t \ddot u_i^2\bigl( q(t') \bigr) \, dt'
  \propto
  \alpha^2 \, \mathcal E.
  $$
  This correction is of order $\alpha^2$
  compared to the main contribution in Eq. \eqref{eq:error_mbin1}.
  %
  Thus, we can assume Eq. \eqref{eq:kappa_delta},
  once $\alpha(t)$ has dropped below a certain level.
  \\
}
\note{
  The extremal property, Eq. \eqref{eq:optimal_mbin},
  now becomes
  %
  \begin{align}
    \sum_{i=1}^n
      \Gamma_i \, \lambda_i \,
      u_i\bigl( q(\tau) \bigr) \,
      \ddot u_i\bigl( q(\tau) \bigr) = 0.
    \label{eq:optimal_mbin1}
  \end{align}
  %
  If we interpret $u_i$ as the position,
  and $\Gamma_i \, \lambda_i$ as the mass,
  then the right-hand side give the virial.
  %
  Thus, the extremal condition demands
  the total virial to be zero.
  $$\,$$
}
%
For example,
if we assume perfect sampling,
%
\begin{equation}
  \langle \zeta_i(t) \, \zeta_j(t') \rangle
  =
  p_i \, (1 - p_i) \, \delta_{ij} \, \delta(t - t').
  \label{eq:zz_perfect}
\end{equation}
%
Thus, from Eqs. \eqref{eq:eig_orthonormal_rows} and \eqref{eq:eta_def},
we get
%
\begin{equation}
  \kappa_i(t - t')
  =
  \sum_{j,k = 1}^n
  \frac{ \phi_{ij} } { p_j }
  \frac{ \phi_{ik} } { p_k }
  \langle \zeta_j(t) \, \zeta_k(t') \rangle
  %
  \approx \delta(t - t'),
  \label{eq:kappa_perfect}
\end{equation}
%
or
$\Gamma_i \approx 1$,
where we have assumed the large $n$ limit
such that $p_i \ll 1$
in the last expression.

%
%We can study two limiting cases.
%
For $\tau \ll t$,
the dominant contribution comes from
the term(s) with the smallest eigenvalue $\lambda_i = \lambda_n$,
and
$$
\ddot u_n(\tau) = 0,
\quad
\dot u_n(\tau) = c,
\quad
\mathrm{and}\;
u_n(\tau) = c \, (\tau + t_0),
$$
which means
$$
\alpha(\tau) \approx \frac{1}{\lambda_n \, (\tau + t_0)}.
$$

@@

We now show
that the single-bin scheme
is an optimal scheme.
%
Consider a class of $\mathbf w$ matrices
with the same set of eigenvectors,
hence same $\Gamma_i$'s,
but with different eigenvalues, $\lambda_i$'s.
%
Using the Cauchy-Schwarz inequality, we have,
for any set of $\beta_i \ge 0$
%
\begin{align}
&
\left( \int_0^t
  dt' \sum_{i = 2}^n \Gamma_i \,
                     \dot u_i^2\bigl( q(t') \bigr)
\right)
\left(
\int_0^t
  dt' \sum_{i = 2}^n \Gamma_i \, c_i^2
\right)
\notag
\\
&
\ge
\left(
\int_0^t
  dt' \sum_{i = 2}^n \Gamma_i \, c_i \,
                     \dot u_i\bigl( q(t') \bigr)
\right)^2
\notag
\\
&
=
\left(
  \sum_{i = 2}^n \Gamma_k \, c_i
    \left[
      1 - e^{ -\lambda_i q(t) }
    \right]
\right)^2.
\notag
\end{align}
%
\note{A short proof of the above inequality
  is the following.
  %
  First, we have
$$
\int_0^t
  dt' \sum_{i = 2}^n \Gamma_i \,
    \left( \dot u_i \, X - c_i \right)^2
\ge 0,
$$
  for any $X$.
  %
  Thus, the left-hand side,
  expanded as a quadratic polynomial of $X$,
  $A \, X^2 + B \, X + C$,
  must have a non-positive discriminant,
  $B^2 - 4 \, A \, C$.
  %
  This condition gives the desired inequality.
  \\%
}
%
Note that the last expression depends on
$q(t')$ only through the end point value,
which is fixed in the variation process.
%
Thus, the inequality sets a lower bound
of the error $\mathcal E$
expressed in Eq. \eqref{eq:error_mbin1},
%
The equality is achieved
if $\dot u_i\left( q(t') \right) = c_i$
for all $i = 2, \dots, n$ at any $t'$,
up to a multiplicative factor.
%
Solving this set of equations yields
$$
\alpha(t') = \frac{              1              }
                  { \lambda_i \, (t' + t_{i0} ) },
\quad
\mathrm{with\;\;}
t_{i0} = \frac{             t            }
              { e^{ \lambda_i q(t) } - 1 }.
$$
Such a solution is possible only if
\begin{equation}
\lambda_2 = \dots = \lambda_n.
\end{equation}
%
\note{
  We add some steps for solution.
  Integrating $\dot u_i = c_i$ yields
  \begin{equation}
  u_i(t') = c_i \left(t' + t_{i0} \right).
  \label{eq:ui_solution}
  \end{equation}
  Given that $u_i(t) = 1$ and $u_i(0) = e^{-\lambda_i \, q(t)}$,
  we have
  $$
  c_i = \frac{ 1 }{ t + t_{i0} },
  \quad
  \mathrm{and\;}
  \frac{ t_{i0} } { t + t_{i0} }
  =
  e^{ -\lambda_k \, q(t) }.
  $$
  Taking the logarithm of Eq. \eqref{eq:ui_solution} yields
  $$
  -\lambda_i \, q(t) + \lambda_i \, q(t')
  = \log c_i + \log\left( t' + t_{i0} \right).
  $$
  Differentiating this with respect to $t'$ yields
  $$
  \lambda_i \, \alpha(t') = \frac{ 1 } { t' + t_{i0} }.
  $$
}

@@

This matrix, being symmetric,
satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
for the flat distribution $p_i \equiv 1/n$.
%
It has $n$ orthonormal eigenvectors,
$\pmb\varphi^{(1)}, \dots, \pmb\varphi^{(n)}$,
given by
%
\begin{equation}
\varphi^{(k)}_i
= \sqrt{
    \frac{ 2 - \delta_{k, 1} }
         {       n           }
       }
  \cos \frac{ \left(i - \frac12\right) (k - 1) \, \pi}{n}.
\label{eq:wband_eigenvector}
\end{equation}
%
with eigenvalues
\begin{align}
  \lambda_k
  &=
  \mu_0 + \sum_{l = 1}^b 2 \, \mu_l \cos \frac{(k - 1)  \, l \pi}{n}
  \notag \\
  &=
  1 - \sum_{l = 1}^b 4 \, \mu_l \sin^2 \frac{(k - 1)  \, l \pi}{ 2 \, n }
  .
\label{eq:wband_eigenvalue}
\end{align}
%
\note{We derive Eqs.
  \eqref{eq:wband_eigenvector} and \eqref{eq:wband_eigenvalue}
  below.

>>>>

To proceed, we first observe that
the average $\bar v = \sum_{i = 1}^n p_i \, v_i$
increases steadily over time:
%
\begin{equation}
\frac{ d \bar v } { d t }
=
\sum_{i = 1}^n p_i \dot v_i
=
\alpha(t) \sum_{i = 1}^n h_i(t) = \alpha(t).
\label{eq:dvbardt}
\end{equation}
%
However, we expect the difference between $v_i$ and $\bar v$,
%
\begin{equation}
  x_i \equiv v_i - \bar v = v_i - \sum_{j = 1}^n p_j \, v_j,
  \label{eq:x_def}
\end{equation}
%
to be small in the asymptotic regime,
and expand Eq. \eqref{eq:h_ave} as
%
\begin{align}
\langle h_i(t) \rangle
&\approx
\frac{                 p_i \, e^{- x_i} }
     { \sum_{ j = 1}^n p_j \, e^{- x_j} }
\approx
\frac{                 p_i (1 - x_i) }
     { \sum_{ j = 1}^n p_j (1 - x_j) }
\notag
\\
&\approx
p_i \, \left(
         1 - x_i + \sum_{j=1}^n p_j \, x_j
       \right)
=
p_i \, (1 - x_i),
\label{eq:hdet_linearize}
\end{align}
%
where we have used $\sum_{j=1}^n p_j = 1$,
and
%
\begin{equation}
  \sum_{i = 1}^n p_i \, x_i = 0,
  \label{eq:px_sum}
\end{equation}
which follows directly from Eq. \eqref{eq:x_def}.
%
From Eqs.
\eqref{eq:vt_diffeq},
\eqref{eq:h_split},
\eqref{eq:dvbardt}-\eqref{eq:hdet_linearize},
we get a set of decoupled equations
%
\begin{equation}
  \dot x_i(t)
  =
  -\alpha(t) \, \left[ x_i(t) - \frac{ \zeta_i(t) } { p_i } \right].
  \label{eq:dxdt_WL}
\end{equation}
\note{The derivation:
$$
\begin{aligned}
  \dot x_i
  &\stackrel{\mathrm{Eq.\; \eqref{eq:x_def}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=}
  \dot v_i - \dot{ \bar v }
  \stackrel{\mathrm{Eqs.\; \eqref{eq:vt_diffeq}\; and\; \eqref{eq:dvbardt}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left( \frac{ h_i } { p_i } - 1 \right)
  \\
  &\stackrel{\mathrm{Eq.\; \eqref{eq:h_split}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left( \frac{ \langle h_i \rangle } { p_i } + \frac{ \zeta_i } { p_i } - 1 \right)
  \stackrel{\mathrm{Eq.\; \eqref{eq:hdet_linearize}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left[ -x_i(t) + \frac{ \zeta_i } { p_i } \right].
\end{aligned}
$$
}

@@

Further, by Eq. \eqref{eq:w_detailedbalance},
we can define a symmetric matrix $\hat{\mathbf w}$
as $\hat w_{ij} = \sqrt{p_i/p_j} \, w_{ij}$
that can be diagonalized\cite{vankampen}
with a set of orthonormal eigenvectors:
%
$\sum_{i = 1}^n \varphi_{ki} \, \hat w_{ij} = \lambda_k \, \varphi_{kj}$.
%
Thus,
in terms of diagonalizing $\mathbf w$, we have
\begin{equation}
  \sum_{i = 1}^n \phi_{ki} \, w_{ij} = \lambda_k \, \phi_{kj},
  \label{eq:eig_w}
\end{equation}
where
$\phi_{ij} \equiv \sqrt{p_j} \, \varphi_{ij}$ satisfies
the orthonormal conditions\cite{vankampen}:
%
\begin{align}
\sum_{k = 1}^n \phi_{ki} \, \phi_{kj}
&= \delta_{ij} \, p_i,
\label{eq:eig_orthonormal_cols}
\\
\sum_{k = 1}^n \frac{ \phi_{ik} \, \phi_{jk} }{ p_k }
&= \delta_{ij}.
\label{eq:eig_orthonormal_rows}
\end{align}

@@

In the asymptotic regime,
we can approximate $\kappa_i(t)$ as
%
\begin{equation}
  \kappa_i(t) \approx \Gamma_i \, \delta(t)
  \label{eq:kappa_delta}
\end{equation}
%
where,
\begin{equation}
  \Gamma_i = \int_{-\infty}^\infty \kappa_i(t) \, dt,
  \notag
  %\label{eq:Gamma_integral}
\end{equation}
or in discrete time,
\begin{equation}
  \Gamma_i = \sum_{t = -\infty}^\infty \kappa_i(t).
  \label{eq:Gamma_sum}
\end{equation}
%
Then, Eq. \eqref{eq:error_mbin} is simplified as
%
\begin{align}
  \mathcal E
  &=
  \int_0^t
  \sum_{i = 2}^n
  \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr) \, dt'.
  \label{eq:error_mbin1}
\end{align}
%
The next order correction to Eq. \eqref{eq:kappa_delta}
would lead to a correction to $\mathcal E$
that is about $\alpha_{\max}^2$ times as large.

\note{This is equivalent to saying that the Fourier transform
  $$
  \tilde \kappa_i(\omega) = \Gamma_i
  $$
  is a roughly constant.
  %
  For the next order correction we would have
  $$
  \tilde \kappa_i(\omega) = \Gamma_i + \Gamma^{(2)}_i \omega^2 + \dots,
  $$
  which leads to a correction to $\mathcal E$
  $$
  \Gamma^{(2)}_i
  \int_0^t \ddot u_i^2\bigl( q(t') \bigr) \, dt'
  \propto
  \alpha^2 \, \mathcal E.
  $$
  This correction is of order $\alpha^2$
  compared to the main contribution in Eq. \eqref{eq:error_mbin1}.
  %
  Thus, we can assume Eq. \eqref{eq:kappa_delta},
  once $\alpha(t)$ has dropped below a certain level.
}

\note{
  The extremal property, Eq. \eqref{eq:optimal_mbin},
  now becomes
  %
  \begin{align}
    \sum_{i=1}^n
      \Gamma_i \, \lambda_i \,
      u_i\bigl( q(\tau) \bigr) \,
      \ddot u_i\bigl( q(\tau) \bigr) = 0.
    \label{eq:optimal_mbin1}
  \end{align}
  %
  If we interpret $u_i$ as the position,
  and $\Gamma_i \, \lambda_i$ as the mass,
  then the right-hand side gives the virial.
  %
  Thus, the extremal condition demands
  the total virial to be zero.
  $$\,$$
}

\subsubsection{Integrals of the correlation functions}

We can evaluate the $\Gamma_i$'s
in two special case.
%
If the sampling is perfect,
Eq. \eqref{eq:kappa_delta} is exact, and
%
\begin{equation}
  \Gamma_i = 1.
  \label{eq:Gamma_perfect}
\end{equation}
%
In this case,
the bin index $i$ is an independent random variable
at each time step $t$, and
%
\begin{equation}
  \langle \zeta_i(t) \, \zeta_j(t') \rangle
  =
  \left( p_i \, \delta_{ij} - p_i \, p_j \right) \, \delta(t - t'),
  \label{eq:zz_perfect}
\end{equation}
%
where $\delta(t - t')$ is understood as an adaption of $\delta_{t, t'}$
from the discrete time case to the continuous limit. % hereinafter.
%
\note{To show Eq. \eqref{eq:zz_perfect}, we first note that
  with $t \ne t'$ (going back to the discrete time case),
  $$
  \begin{aligned}
  \langle \zeta_i(t) \, \zeta_j(t') \rangle
  &=
  \langle \zeta_i(t)  \rangle
  \langle \zeta_j(t') \rangle
  \\
  &=
  \bigl\langle h_i(t)  - \langle h_i(t)  \rangle \bigr\rangle
  \bigl\langle h_j(t') - \langle h_j(t') \rangle \bigr\rangle
  =
  0.
  \end{aligned}
  $$
  Next, for $t = t'$, we have
  $$
  \zeta_i(t)
  = h_i(t) - \langle h_i(t) \rangle
  = \delta_{i, i(t)} - p_i,
  $$
  and
  $$
  \begin{aligned}
  \left\langle
    \zeta_i(t) \, \zeta_j(t)
  \right\rangle
  &=
  \left\langle
    \delta_{i, i(t)} \, \delta_{j, i(t)}
  \right\rangle
  -
  p_i \, p_j
  \\
  &=
  \left\langle
    \delta_{i, i(t)}
  \right\rangle
  \delta_{i, j}
  -
  p_i \, p_j
  =
  p_i \, \delta_{ij}
  - p_i \, p_j.
  \end{aligned}
  $$
  %
  The average
  $\left\langle \delta_{i, i(t)} \right\rangle$ is $p_i$,
  because as $i(t)$ goes through $1$ to $n$,
  $\delta_{i, i(t)}$ gives $1$
  if $i(t)$ matches $i$ with probability $p_i$,
  or $0$ otherwise.
}
%
Thus, from Eqs. \eqref{eq:eig_orthonormal_rows} and \eqref{eq:eta_def},
we get, for $i \ge 2$,
%
\begin{equation}
  \kappa_i(t - t')
  =
  \sum_{j,k = 1}^n
  \frac{ \phi_{ij} } { p_j }
  \frac{ \phi_{ik} } { p_k }
  \langle \zeta_j(t) \, \zeta_k(t') \rangle
  %
  = \delta(t - t'),
  \label{eq:kappa_perfect}
\end{equation}
%
\note{To show Eq. \eqref{eq:kappa_perfect},
  $$
  \begin{aligned}
  \kappa_i(t - t')
  &=
  \sum_{j,k = 1}^n
  \frac{ \phi_{ij} } { p_j }
  \frac{ \phi_{ik} } { p_k }
  \langle \zeta_j(t) \, \zeta_k(t') \rangle
  \\
  %
  &=
  \sum_{j,k = 1}^n
  \frac{ \phi_{ij} } { p_j }
  \frac{ \phi_{ik} } { p_k }
  \left(
    p_j \, \delta_{j, k} - p_j \, p_k
  \right)
  \, \delta(t - t')
  \\
  &=
  \left(
    \sum_{j = 1}^n
    \frac{ \phi_{ij}^2 } { p_j }
    -
    \sum_{j = 1}^n \phi_{ij}
    \sum_{k = 1}^n \phi_{ik}
  \right)
  \, \delta(t - t')
  \\
  &=
  (1 - 0 \cdot 0) \, \delta(t - t').
  \end{aligned}
  $$
  where,
  we have used
  Eq. \eqref{eq:eta_def}
  on the first line,
  %
  Eq. \eqref{eq:zz_perfect}
  on the second line,
  and
  Eq. \eqref{eq:eig_orthonormal_rows}
  on the last line.
  %
  Note for $i \ge 2$,
  $$
  \sum_{j = 1}^n \phi_{ij}
  =
  \sum_{j = 1}^n \frac{ \phi_{ij} \, \phi_{1j} } { p_j }
  =
  \delta_{i1}
  =
  0.
  $$
  as the mode $1$, or the first row, $\phi_{1j}$,
  is given by the equilibrium distribution $p_j$.
}

Another special case is the one-step process\cite{vankampen},
in each step, the system hops to the previous $i - 1$ bin
or the next bin $i + 1$ with equal probability $1/2$,
ignoring out-of-boundary transitions.
%
We shall further assume that
the eigenvectors of the $\mathbf w$ matrix
coincide with the those of the transition matrix
of the above sampling process
(cf. Sec. \ref{sec:band-matrix}).
%
Since the transition matrix is tridiagonal,
as given by Eq. \eqref{eq:wtridiag} with $\mu_1 = 1/2$,
the eigenvalues are given
by Eq. \eqref{eq:wband_eigenvalue} as
$$
\Lambda_i = 1 - 2 \sin^2 \frac{ (i - 1) \, \pi } { 2 \, n }.
$$
Note, however, these $\Lambda_i$'s are the eigenvalues of
the transition matrix of sampling scheme,
not to be confused with the eigenvalues, $\lambda_i$'s,
of the $\mathbf w$ matrix.
%
The correlation function $\kappa_i(t)$ of
decays as $\kappa_i(0) \, \Lambda_i^t$,
with the same-time value $\kappa_i(0) = 1$ being the same as
that in perfect sampling.
%
Thus, from Eq. \eqref{eq:Gamma_sum}, we get
%
\begin{equation}
\Gamma_i
=
1 + 2 \, \sum_{t = 1}^\infty \Lambda_i^t
=
\frac{ 1 + \Lambda_i } { 1 - \Lambda_i }
=
\cot^2 \left( \frac{ i \, \pi } { 2 \, n } \right).
\label{eq:Gamma_onestep}
\end{equation}
%
We can use this case to model local sampling processes.

%
%We can study two limiting cases.
%
%For $\tau \ll t$,
%the dominant contribution comes from
%the term(s) with the smallest eigenvalue $\lambda_i = \lambda_n$,
%and
%$$
%\ddot u_n(\tau) = 0,
%\quad
%\dot u_n(\tau) = c,
%\quad
%\mathrm{and}\;
%u_n(\tau) = c \, (\tau + t_0),
%$$
%which means
%$$
%\alpha(\tau) \approx \frac{1}{\lambda_n \, (\tau + t_0)}.
%$$

@@

We now show
that the single-bin scheme
is an optimal scheme.
%
Consider a class of $\mathbf w$ matrices
sharing the same set of eigenvectors,
hence same $\Gamma_i$'s,
but with different eigenvalues, $\lambda_i$'s.
%
Using the Cauchy-Schwarz inequality, we have,
for any set of $\beta_i \ge 0$
%
\begin{align}
&
\left(
  \int_0^t dt'
    \sum_{i = 2}^n
      \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr)
\right)
%
\left(
  \int_0^t dt'
    \sum_{i = 2}^n
      \Gamma_i \, c_i^2
\right)
%
\notag
\\
&
\ge
\left(
  \int_0^t dt'
    \sum_{i = 2}^n
      \Gamma_i \, c_i \, \dot u_i\bigl( q(t') \bigr)
\right)^2
\notag
\\
&
=
\left(
  \sum_{i = 2}^n \Gamma_k \, c_i
    \left[
      1 - e^{ -\lambda_i q(t) }
    \right]
\right)^2.
\notag
\end{align}
%
\note{A short proof of the above inequality
  is the following.
  %
  First, we have
$$
\int_0^t
  dt' \sum_{i = 2}^n \Gamma_i \,
    \left( \dot u_i \, X - c_i \right)^2
\ge 0,
$$
  for any $X$.
  %
  Thus, the left-hand side,
  expanded as a quadratic polynomial of $X$,
  $A \, X^2 + B \, X + C$,
  must have a non-positive discriminant,
  $B^2 - 4 \, A \, C$.
  %
  This condition gives the desired inequality.
  \\%
}
%
Note that the last expression depends on
$q(t')$ only through the end point value,
which is fixed in the variation process.
%
Thus, the inequality sets a lower bound
of the error $\mathcal E$
expressed in Eq. \eqref{eq:error_mbin1},
%
The equality is achieved
if $\dot u_i\left( q(t') \right) = c_i$
for all $i = 2, \dots, n$ at any $t'$,
up to a multiplicative factor.
%
Solving this set of equations yields
$$
\alpha(t') = \frac{              1              }
                  { \lambda_i \, (t' + t_{i0} ) },
\quad
\mathrm{with\;\;}
t_{i0} = \frac{             t            }
              { e^{ \lambda_i q(t) } - 1 }.
$$
Such a solution is possible only if
\begin{equation}
\lambda_2 = \dots = \lambda_n.
\end{equation}
%
\note{
  We add some steps for solution.
  Integrating $\dot u_i = c_i$ yields
  \begin{equation}
  u_i(t') = c_i \left(t' + t_{i0} \right).
  \label{eq:ui_solution}
  \end{equation}
  Given that $u_i(t) = 1$ and $u_i(0) = e^{-\lambda_i \, q(t)}$,
  we have
  $$
  c_i = \frac{ 1 }{ t + t_{i0} },
  \quad
  \mathrm{and\;}
  \frac{ t_{i0} } { t + t_{i0} }
  =
  e^{ -\lambda_k \, q(t) }.
  $$
  Taking the logarithm of Eq. \eqref{eq:ui_solution} yields
  $$
  -\lambda_i \, q(t) + \lambda_i \, q(t')
  = \log c_i + \log\left( t' + t_{i0} \right).
  $$
  Differentiating this with respect to $t'$ yields
  $$
  \lambda_i \, \alpha(t') = \frac{ 1 } { t' + t_{i0} }.
  $$
}

@@

This matrix, being symmetric,
satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
for the flat distribution $p_i \equiv 1/n$.
%
It has $n$ orthonormal eigenvectors,
$\pmb\varphi^{(1)}, \dots, \pmb\varphi^{(n)}$,
given by
%
\begin{equation}
\varphi^{(k)}_i
= \sqrt{
    \frac{ 2 - \delta_{k, 1} }
         {       n           }
       }
  \cos \frac{ \left(i - \frac12\right) (k - 1) \, \pi}{n}.
\label{eq:wband_eigenvector}
\end{equation}
%
with eigenvalues
%
\begin{align}
  \lambda_k
  &=
  \mu_0 + \sum_{l = 1}^b 2 \, \mu_l \cos \frac{(k - 1)  \, l \pi}{n}
  \notag \\
  &=
  1 - \sum_{l = 1}^b 4 \, \mu_l \sin^2 \frac{(k - 1)  \, l \pi}{ 2 \, n }
  .
\label{eq:wband_eigenvalue}
\end{align}
%
\note{We derive Eqs.
  \eqref{eq:wband_eigenvector} and \eqref{eq:wband_eigenvalue}
  below.
